Modal Abstractions for Virtualizing Memory Addresses

### 2023-11-15 examples.tex
+

Most kernels maintain an invariant on their page tables that the virtual address of any page used for a page table lives at a virtual address
whose value is a constant offset from the physical address a practice sometimes referred to as \emph{identity mapping} (even though the physical-to-virtual translation
is typically not literally the identity function, but adding a constant offset).\footnote{Some kernels do this for all physical memory on the machine, simplifying interaction
with DMA devices.}

@@

\todo[inline]{typeset as a triple}

>>>>

Most kernels maintain an invariant on their page tables that the virtual address of any page used for a page table lives at a virtual address
whose value is a constant offset from the physical address a practice sometimes referred to as \emph{identity mapping} (even though the physical-to-virtual translation
is typically not literally the identity function, but adding a constant offset).\footnote{Some kernels do this for all physical memory on the machine, simplifying interaction
with DMA devices.}
\newcommand{\qfraczero}{\textsf{qfrac}}
\newcommand{\true}{\textsf{true}}
\tikzstyle{boxedassert_border} = [sharp corners,line width=0.2pt]
\NewDocumentCommand \boxedassertpv {O{} m o}{%
	\tikz[baseline=(m.base)]{
		%	  \node[rectangle, draw,inner sep=0.8pt,anchor=base,#1] (m) {${#2}\mathstrut$};
		\node[rectangle,inner sep=1.5pt,outer sep=0.2pt,anchor=base] (m) {${\,#2\,}\mathstrut$};
		\draw[#1,boxedassert_border] ($(m.south west)$) rectangle ($(m.north east)$);
	}\IfNoValueF{#3}{^{\,#3}}%
}
\newcommand*{\knowInvpv}[2]{\boxedassertpv{#2}[#1]}
\newcommand*{\ownGhostpv}[2]{\boxedassertpv[dash dot]{#2}[#1]}

\newcommand{\sumpv}[3]{
  \ownGhostpv\gammaPred{\authfrag{\singletonMap{#1}{(#2, #3)}}}
}

\newcommand{\pvmapping}[1]{\mathcal{A}\textsf{PVMappings}(#1)}
\newcommand{\sumapacesfull}[2]{
  \ownGhost\gammaPreds{\authfull{\singletonMap{#1}{#2}}}
}
\begin{figure*}
\[
\begin{array}{l}
  \mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)\stackrel{\triangle}{=} \textsf{ASpace\_Lookup}(\ptablestore,\Xi,m) \ast \\
  \qquad\qquad\qquad \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})} \\
  \qquad\qquad\qquad \bigast{(\paddr) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{KERNELBASE} \urcorner \ast  \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{Ghost translation} \ast \\
\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad    \underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\}\; \vale}_\text{Physical location} \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner }\\
  \textsf{ where } \\
   \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \exists \gammaPred \; \ldotp \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammaPred \urcorner \ast
   \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \\ \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad \ownGhostpv\gammaPred{\authfull{\pvmapping\Xi}}

\end{array}
\]
\caption{Global Address-Space Invariant in Figure \ref{fig:peraspaceinvariant} extended with a ghost map bookkeeping identity mappings }
  \label{fig:peraspaceinvariant_with_p2v_extension}
  \end{figure*}

@@

\begin{figure}\footnotesize
\begin{lstlisting}[mathescape]
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner \textsf{KERNELBASE} + \paddr > 0 \urcorner }_{\rtv}$
$\specline{ \texttt{rip}\mapsto_{\textsf{r}} \textsf{iv} \ast \texttt{rflags}\mapsto_{\textsf{r}} \textsf{flaglist} \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr  }_{\rtv}$
$\specline{ \sumapaces\rtv\delta  \ast \sumwalkabs\vaddr\qfrac\fpaddr \ast \sumpv\fpaddr\qfrac\true  }_{\rtv}$
;; This offsetst physical address at KERNELBASE!
add rbx, KERNELBASE
$\specline{\exists \qfraczero; \qfrac,\textsf{ flaglistadd, w, flaglistadd, val}, \textsf{P} \ast \texttt{rip}\mapsto_{\textsf{r}} \textsf{iv+7} \ast \texttt{rflags}\mapsto_{\textsf{r}} \textsf{flaglistadd} \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr + \textsf{KERNELBASE} }_{\rtv}$
$\specline{\bigast{(\vaddr, \paddr)\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, } \paddr)\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e,} \paddr)}}_{\rtv}$
$\specline{\mathcal{R} -\ast \bigast{(\paddr) \in \Xi}{\exists\; (\textsf{qfrac, q, val, va}) \ldotp \ulcorner \textsf{va} = \paddr + \textsf{KERNELBASE} \urcorner \ast \sumwalkabs\vaddr\qfrac\paddr  \ast \paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\} \; \vale\ast  \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner}}_{\rtv}$
$\specline{ \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \ownGhostpv\gammaPred{\authfull{\pvmapping\Xi}}}$
$\specline{ \sumapaces\rtv\delta  \ast\sumwalkabs\vaddr\qfrac\paddr \ast \sumpv\paddr\qfraczero\true \ast \textsf{P}}_{\rtv}$
$\specline{\vaddr \mapsto_{\texttt{vpte}} \{\textsf{q}, \textsf{qfrac}\} \paddr \;(\texttt{val})}$
\end{lstlisting}
\caption{Identity mapping of a physical address to a virtual one.}
\label{fig:mapping_code}
\end{figure}
